%% Taught MSc -- specify a particular degree instead. If none is specified,
%% "MSc in Informatics" is used.
\documentclass[msc,cs,twoside,openright,logo,rightchapter,normalheadings]{class/infthesis}
% \documentclass[msc]{infthesis}  % for the MSc in Informatics

%% Packages
\usepackage[utf8]{inputenc}   % Enable UTF-8 typing
\usepackage[british]{babel}   % British English
\usepackage[breaklinks]{hyperref}         % Interactive PDF
\usepackage{url}
\usepackage{breakurl}
\usepackage{amsmath}          % Mathematics library
\usepackage{amssymb}          % Provides math fonts
\usepackage{amsthm}           % Provides \newtheorem, \theoremstyle, etc.
\usepackage{mathpartir}       % Inference rules
\usepackage{array}
\usepackage{float}            % Float control
\usepackage{caption,subcaption}  % Sub figures support
\usepackage[T1]{fontenc}      % Fixes font issues
\usepackage{lmodern}

\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}  
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\usepackage[
backend=bibtex,
style=numeric,
sorting=nyt
]{biblatex}                   % Bibliography
\addbibresource{references.bib}

% Source code listings
\usepackage{xcolor}
\usepackage{listings}

\lstset{
 backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
 basicstyle=\ttfamily\small,        % the size of the fonts that are used for the code
 commentstyle=\itshape,
 breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
 breaklines=true,                 % sets automatic line breaking
 captionpos=b,                    % sets the caption-position to bottom
 deletekeywords={...},            % if you want to delete keywords from the given language
 escapeinside={§*}{*§},          % if you want to add LaTeX within your code
 extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
 frame=none,	                   % adds a frame around the code
 keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
 numbers=none,                    % where to put the line-numbers; possible values are (none, left, right)
 rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
 showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
 showstringspaces=false,          % underline spaces within strings only
 showtabs=false,                  % show tabs within strings adding particular underscores
 tabsize=2,	                   % sets default tabsize to 2 spaces
 title=\lstname,                   % show the filename of files included with \lstinputlisting; also try caption instead of title
  belowcaptionskip=-1\baselineskip,
  xleftmargin=\parindent
}
% Haskell style
\lstdefinestyle{haskell}{
  language=Haskell,
  basicstyle=\linespread{1.0}\ttfamily\footnotesize,
  literate= {+}{{$+$}}1 {*}{{$*$}}1
            {<=}{{$\leq$}}1 {/=}{{$\neq$}}1 
            {==}{{$\equiv$}}1 {=>}{{$\Rightarrow$}}1
            {->}{{$\to$}}1 {<-}{{$\leftarrow$}}1
            {.}{{$\circ$}}1 {$$}{{\$}}1
}
% Ocaml style
\lstdefinestyle{ocaml}{
  language=Caml,
  literate= {+}{{$+$}}1 {*}{{$*$}}1
            {<=}{{$\leq$}}1 {>=}{{$\geq$}}1 {<>}{{$\neq$}}1 
            {==}{{$\equiv$}}1 {=>}{{$\Rightarrow$}}1
            {->}{{$\to$}}1
}
% Links style
\lstdefinestyle{links}{
  basicstyle=\linespread{1.0}\ttfamily\footnotesize,
  language=Caml,
  literate= {+}{{$+$}}1 {*}{{$*$}}1
            {<=}{{$\leq$}}1 {>=}{{$\geq$}}1 
            {=>}{{$\Rightarrow$}}1
            {->}{{$\to$}}1 {~>}{{$\rightsquigarrow$}}1
}

% Example environment
\makeatletter  
\def\@endtheorem{\qed\endtrivlist\@endpefalse } % insert `\qed` macro
\makeatother
\theoremstyle{definition}
\newtheorem{example}{Example}[chapter]

% Convenient macros
\newcommand{\record}[1]{\ensuremath{\left\{ #1 \right\}}} %{\langle #1 \rangle}}
\newcommand{\typename}[1]{\texttt{#1}}
\newcommand{\type}[1]{\typename{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\eto}[1]{\ensuremath{\xrightarrow{\left\{#1 \right\}\;}}}
\newcommand{\thunktype}[2]{\ensuremath{() \eto{#1} #2}}
\newcommand{\chntype}[2]{\ensuremath{(#1) \to #2}}
\newcommand{\defas}[0]{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}} % "defined-as-equal"
\newcommand{\rulesep}{\unskip\ \vrule\ } % Inserts a vertical line

%% Information about the title, etc.
\title{Handlers for Algebraic Effects in Links}
\author{Daniel Hillerström}

%% If the year of submission is not the current year, uncomment this line and 
%% specify it here:
\submityear{2015}

%% Optionally, specify the graduation month and year:
% \graduationdate{February 1786}

%% Specify the abstract here.
\abstract{%
Plotkin and Pretnar's handlers for algebraic effects provide a compelling alternative to monads as a basis for effectful programming. In their programming model computations are composed from abstract operations. Handlers interpret abstract operations by instantiating them with concrete implementations. Consequently, the model promises a high degree of modularity.

We provide an implementation of handlers for algebraic effects using \emph{row polymorphism} in the strict functional programming language Links. Row polymorphism enable us to truly compose computations from operations in any order.

Through a series of examples we demonstrate that the handler abstraction makes it easy to write modular, effectful programs. Furthermore, we show that the compositionality of handlers enable of us to extend the functionality of effectful programs effortlessly.

Finally, we also include a discussion of our implementation as well as an evaluation of our programming model, how handlers and effects fit into the Links language, and an account of the performance of handlers.
}

%% Now we start with the actual document.
\begin{document}
\raggedbottom
%% First, the preliminary pages
\begin{preliminary}

%% This creates the title page
\maketitle

%% Acknowledgements
\begin{acknowledgements}
I would like to extend my sincere gratitude to my supervisor \href{http://homepages.inf.ed.ac.uk/slindley/}{Sam Lindley}. Firstly, for accepting me as his student, and secondly, for introducing to the subject. Sam has provided invaluable guidance during the entire project period. We have enjoyed many conversations about programming language theory over the summer.

I would also like to thank \href{https://www.cs.st-andrews.ac.uk/directory/person?id=eb}{Edwin Brady} for inviting me to attend and deliver a talk about my work during the \emph{Practical Types 2015}, summer school at the University of St. Andrews. It gave me an opportunity to receive some valuable feedback on my work.

Finally, thanks to \href{http://homepages.inf.ed.ac.uk/jrl/}{John Longley}, \href{http://homepages.inf.ed.ac.uk/dcspaul/homepage/live/work/index.html}{Paul Anderson}, and \href{http://homepages.inf.ed.ac.uk/slindley/}{Sam Lindley} for their feedback during the midterm MSc progress seminar.
\end{acknowledgements}

%% Next we need to have the declaration.
\standarddeclaration

%% Finally, a dedication (this is optional -- uncomment the following line if
%% you want one).
%\dedication{To my mummy.}

\begin{preface}
This report contains a variety of examples. Some of the examples elide aesthetic details to condense the code. The complete source code for every example, including instructions, is available at \url{https://github.com/dhil/msc-dissertation}. The repository also contains additional examples that are not presented in this report. Furthermore, the source code for the handlers implementation in Links is available at \url{https://github.com/dhil/links-effect-handlers}.
\end{preface}

%% Create the table of contents
\setcounter{secnumdepth}{2} % Numbering on sections and subsections
\setcounter{tocdepth}{1} % Show only chapters and sections
%\singlespace
\tableofcontents
%\doublespace

%% If you want a list of figures or tables, uncomment the appropriate line(s)
% \listoffigures
% \listoftables

\end{preliminary}

%%%%%%%%
%% Include your chapter files here. See the sample chapter file for the basic
%% format.

\input{content/index}

%% If you want the bibliography single-spaced (which is allowed), uncomment
%% the next line.
\singlespace
\printbibliography[heading=bibintoc]

%% ... that's all, folks!
\end{document}
